---
title: "01_EDA"
author: "Sonal Mendiratta"
date: "5/14/2020"
output: html_document
---

# Objective:
* Understanding the data
* Visualization

```{r include = FALSE}
setwd("/Users/sonal/Desktop/Desktop_v2/MSBA/COURSES/QUARTER4/PredictiveAnalytics/Project")
```

```{r include = FALSE}
library(lattice)
library(ggplot2)
library(dplyr)  
library("psych")
library(ggcorrplot)
library(leaps)
library(corrplot)
library(glmnet)
library(boot)
library(glmnet)
library(fpp2)
library(lubridate)
library(sqldf)
options(scipen=999)
library(treemap)
library(leaflet)
library(forecast)
```


#  Understanding the data

### 1.air_reserve 

* air_store_id - the restaurant's id in the air system
* visit_datetime - the time of the reservation
* reserve_datetime - the time the reservation was made
* reserve_visitors - the number of visitors for that reservation

```{r}
air_reserve <- data.frame(read.csv("air_reserve.csv"))
head(air_reserve)
dim(air_reserve)
names(air_reserve)
air_reserve %>% distinct(air_store_id) %>% nrow()
```

314 unique stores in reservation data

### 2. air_store_info
Information around air restaurants

```{r}
air_store_info <- data.frame(read.csv("air_store_info.csv"))
head(air_store_info)
dim(air_store_info)
air_store_info %>% distinct(air_store_id) %>% nrow()
```
829 unique stores 


### 3. air_visit_data
Visit data for air restaurants

```{r}
air_visit_data  <- data.frame(read.csv("air_visit_data.csv"))
head(air_visit_data)
dim(air_visit_data)
names(air_visit_data)
air_visit_data %>% distinct(air_store_id) %>% nrow()
```
829 unique stores 

### 4. date_info
calendar date information - day of the week and holiday
```{r}
date_info <- data.frame(read.csv("date_info.csv"))
head(date_info)
dim(date_info)
str(date_info)
date_info %>% distinct(calendar_date) %>% nrow()
```
 517 unique dates

### 5. hpg_reserve
details of reservations made in the hpg system
```{r}
hpg_reserve <- data.frame(read.csv("hpg_reserve.csv"))
head(hpg_reserve)
dim(hpg_reserve)
hpg_reserve %>% distinct(hpg_store_id) %>% nrow()
```
13325 distinct stores 

###6. hpg_store_info
information around hpg restaurants
```{r}
hpg_store_info <- data.frame(read.csv("hpg_store_info.csv"))
head(hpg_store_info)
dim(hpg_store_info)
hpg_store_info %>% distinct(hpg_store_id) %>% nrow()
```
4690 unique stores

**Store information is for 892 unique Air stores and 13325 unique hpg stores**

###7. store_id_relation
relation between hpg and air store ids
```{r}
store_id_relation <- data.frame(read.csv("store_id_relation.csv"))
head(store_id_relation)
dim(store_id_relation)
store_id_relation %>% distinct(air_store_id) %>% nrow()
```
relation between hpg id and air id for 150 stores

###8. sample_submission

* final data - need visitor count for each restaurant
* dependent variable -  Visitors
* we are not predicting number of visits in a day but number of visitors in a day!

```{r}
sample_submission <- data.frame(read.csv("sample_submission.csv"))
head(sample_submission)
dim(sample_submission)
```

# Visualization

* distribution of restaurancts by genre
* distribution of restaurants by area
* distribution of visits by holidays
* distribution of visits during the week
* distribution of visits overall

## Data Pre-processing

```{r}
# Converting date from factor to date format in all datasets

# air_visit_data
air_visit_data$visit_date <- ymd(air_visit_data$visit_date)
head(air_visit_data)

# air_reserve
air_reserve$reserve_datetime <- ymd_hms(air_reserve$reserve_datetime)
air_reserve$visit_datetime <- ymd_hms(air_reserve$visit_datetime)

# hpg_reserve
hpg_reserve$reserve_datetime <- ymd_hms(hpg_reserve$reserve_datetime)
hpg_reserve$visit_datetime <- ymd_hms(hpg_reserve$visit_datetime)
head(hpg_reserve)

# date_info
date_info$calendar_date <- ymd(date_info$calendar_date)
date_info$day_of_week <- weekdays(date_info$calendar_date)
```

## AIR Restaurants

```{r}
air_visit_store <- sqldf("select a.*, 
                                 b.air_genre_name,
                                 b.air_area_name,
                                 b.latitude,
                                 b.longitude,
                                 c.day_of_week,
                                 c.holiday_flg
                         from air_visit_data a
                         left join air_store_info b
                         on a.air_store_id = b.air_store_id
                         left join date_info c
                         on a.visit_date = c.calendar_date")
```

Using air_visit_data here as it as data for 829 unique stores 

## number of restaurants by genre
```{r}
air_visit_store %>%                    
  group_by(air_genre_name) %>%         
  summarize(stores = n_distinct(air_store_id)) %>%
  ggplot(aes(x = air_genre_name,  y = stores, fill = air_genre_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(x = "Genre", y = "Number of air restaurants")
```

Majority of the restaurants are of Izakya and Cafe/Sweets genre

## number of restaurants by area
```{r}
air_visit_store %>%                    
  group_by(air_area_name) %>%         
  summarize(stores = n_distinct(air_store_id)) %>%
  treemap(index="air_area_name",
        vSize="stores",
        type="index")
```

Tip 3 populated areas are: Fukuo.., Tokyo.. Tokyo..

## Day of the week
```{r}
air_visit_store %>%
  group_by(day_of_week) %>%
  summarise(visits = sum(visitors)) %>%
  ggplot(aes(day_of_week, visits, fill = day_of_week)) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Day", y = "Total number of visitors") + 
  ggtitle("Distribution of Visits by the day of the week")
```
 
 Highest number of visits are on Saturday followed by Friday

##  Genre
```{r}
air_visit_store %>%
  group_by(air_genre_name) %>%
  summarise(visits = sum(visitors)) %>%
  ggplot(aes(air_genre_name, visits, fill = air_genre_name)) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Genre", y = "Total number of visitors") + 
  ggtitle("Distribution of Visits by Genre")+
  coord_flip()
```

Majority of the visits are also in the same top 2 genres

## Holidays
```{r}
test <- air_visit_store
test$holiday_flg <- as.factor(test$holiday_flg)
ggplot(test,aes(x = holiday_flg,y = visitors, fill = holiday_flg, group = holiday_flg))+
  geom_violin(trim=FALSE, alpha = 0.7) +
  theme(legend.position="none")+
  ggtitle("Distribution of Visits by Holidays")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_boxplot(width=0.1,fill="white")+
  scale_fill_manual(values=c("#F2359D", "#034C8C")) +
  xlab("Holiday")
```

Interestingly, holidays do not really have an impact on the visitors.

## Holidays on weekdays 

```{r}
test <- air_visit_store
test$holiday_flg <- as.factor(test$holiday_flg)
test %>%
  group_by(day_of_week, holiday_flg) %>%
  summarise(mean_visitors = mean(visitors)) %>%
  ggplot(aes(x=day_of_week, y=mean_visitors, color=holiday_flg)) + 
    geom_point(size=4) +
  xlab("Day")+
  ylab("Average number of visitors")+
  theme_light()
```
 More number of visits on a weekday if it is a holiday

## Time series

```{r}
air_vis <- subset(air_visit_store, select = c("visit_date","visitors"))
air_vis1 <- air_vis %>%                    
  group_by(visit_date) %>%         
  summarize(visits = sum(visitors)) 

air_vis_ts <- ts(air_vis1[,2], frequency = 7)
autoplot(air_vis_ts) + xlab("Date") + ylab("Visitors") + ggtitle("Visitors throughout the time frame")
```

```{r}
auto <- auto.arima(air_vis_ts)
summary(auto)
autoplot(auto)
```

```{r}
air_visit_store1 <- sqldf("select distinct air_store_id,
                      longitude,latitude ,air_genre_name
                      from air_visit_store
                          ")
leaflet(air_visit_store1) %>%
  addTiles() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(~longitude, ~latitude,
             popup = ~air_store_id, label = ~air_genre_name,
             clusterOptions = markerClusterOptions())
```

## HPG Restaurants

```{r}
head(hpg_reserve)
head(hpg_store_info)
```


```{r}
#converting the datetime to a date so that i can merge with calendar date
hpg_reserve$visit_date <- as.Date(hpg_reserve$visit_datetime)
```


```{r}
hpg_visit_store <- sqldf("select a.*, 
                                 hpg_genre_name,
                                 hpg_area_name,
                                 b.latitude,
                                 b.longitude,
                                 c.day_of_week,
                                 c.holiday_flg
                         from hpg_reserve a
                         left join hpg_store_info b
                         on a.hpg_store_id = b.hpg_store_id
                         left join date_info c
                on  a.visit_date = c.calendar_date")
```


## number of restaurants by genre - had to remove NAs while plotting as some restraurants do not have the information
```{r}
hpg_visit_store %>%  
   filter(!is.na(hpg_genre_name)) %>%
  group_by(hpg_genre_name) %>%         
  summarize(stores = n_distinct(hpg_store_id)) %>%
  ggplot(aes(x = hpg_genre_name,  y = stores, fill = hpg_genre_name)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(x = "Genre", y = "Number of hpg restaurants")
```

International cuisine is second most common!

## number of restaurants by area
```{r}
hpg_visit_store %>%                    
  group_by(hpg_area_name) %>%         
  summarize(stores = n_distinct(hpg_store_id)) %>%
  treemap(index="hpg_area_name",
        vSize="stores",
        type="index")
```

```{r}
head(hpg_visit_store)
```

## Day of the week
```{r}
hpg_visit_store %>%
  group_by(day_of_week) %>%
  summarise(visits = sum(reserve_visitors)) %>%
  ggplot(aes(day_of_week, visits, fill = day_of_week)) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Day", y = "Total number of visitors") + 
  ggtitle("Distribution of Visits by the day of the week")
```

HPG restaurants are more popular on Friday instead.

##number of visits by genre - had to remove NAs while plotting as some restraurants do not have the information
```{r}
hpg_visit_store %>%
     filter(!is.na(hpg_genre_name)) %>%
  group_by(hpg_genre_name) %>%
  summarise(visits = sum(reserve_visitors)) %>%
  ggplot(aes(hpg_genre_name, visits, fill = hpg_genre_name)) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Genre", y = "Total number of visitors") + 
  ggtitle("Distribution of Visits by Genre")+
  coord_flip()
```

Japanese style and International Cuisine.

## Holidays
```{r}
test <- hpg_visit_store
test$holiday_flg <- as.factor(test$holiday_flg)
ggplot(test,aes(x = holiday_flg,y = reserve_visitors, fill = holiday_flg, group = holiday_flg))+
  geom_violin(trim=FALSE, alpha = 0.7) +
  theme(legend.position="none")+
  ggtitle("Distribution of Visits by Holidays")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_boxplot(width=0.1,fill="white")+
  scale_fill_manual(values=c("#F2359D", "#034C8C")) +
  xlab("Holiday")
```

## Maybe mean visitors is better for visualization

```{r}
test <- hpg_visit_store
test$holiday_flg <- as.factor(test$holiday_flg)
test %>%
  group_by( holiday_flg) %>%
  summarise(mean_visitors = mean(reserve_visitors)) %>%
  ggplot(aes(x=holiday_flg, y=mean_visitors, color=holiday_flg)) + 
    geom_point(size=4) +
  xlab("Holiday")+
  ylab("Average number of visitors")+
  theme_light()
```

On an average, visits are more on a non holiday than a holiday
## Holidays on weekdays 

```{r}
test <- hpg_visit_store
test$holiday_flg <- as.factor(test$holiday_flg)
test %>%
  group_by(day_of_week, holiday_flg) %>%
  summarise(mean_visitors = mean(reserve_visitors)) %>%
  ggplot(aes(x=day_of_week, y=mean_visitors, color=holiday_flg)) + 
    geom_point(size=4) +
  xlab("Day")+
  ylab("Average number of visitors")+
  theme_light()
```

Unless holiday is on weekday..


```{r}
##R studio crases when I run this
# hpg_visit_store1 <-
#   hpg_visit_store %>%
# filter(!is.na(hpg_genre_name))
# 
# leaflet(hpg_visit_store1) %>%
#   addTiles() %>%
#   addProviderTiles("CartoDB.Positron") %>%
#   addMarkers(~longitude, ~latitude,
#              popup = ~hpg_store_id, label = ~hpg_genre_name,
#              clusterOptions = markerClusterOptions())
```

## Time series

```{r}
hpg_vis <- subset(hpg_visit_store, select = c("visit_date","reserve_visitors"))
hpg_vis1 <- hpg_vis %>%                    
  group_by(visit_date) %>%         
  summarize(visits = sum(reserve_visitors)) 

hpg_vis_ts <- ts(hpg_vis1[,2], frequency = 7)
autoplot(hpg_vis_ts) + xlab("Date") + ylab("Visitors") + ggtitle("Visitors throughout the time frame")
```

```{r}
auto <- auto.arima(hpg_vis_ts)
summary(auto)
autoplot(auto)
```

